Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-03-28 Response_Letter/Response_Letter.tex

%%% REVIEWER 1 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#1}

{The manuscript reports neutral excitation energy calculations with the Bethe-Salpeter equation in the $T$-matrix approximation. 
The paper is well written and reports important new results.

However general conclusions like "BSE@GT performs best for few-electron systems whereas BSE@GW is better for larger systems" 
are drawn out of only 3 systems: H2, H2O and BeH2. None of the 3 can be considered as a "large" system.

I think the work would be much stronger with a more comprehensive benchmark set.}
\\
\alert{
We thanks the reviewer for these positive comments.
We are planning on performing a comprehensive benchmark in a forthcoming paper but we feel that it is currently outside the scope of the present paper which is dedicated to the introduction of a new theory and test calculations.
Accordingly with the reviewer's comment, we have soften our conclusion about the accuracy of the BSE@GT for large systems while mentioning that a more comprehensive benchmark study would be required:
``The overall accuracy of the present scheme still needs to be assessed for larger systems (where the screening is known to be more important). 
For such purposes, a comprehensive benchmark study would be required and we are planning to do so in the future''
}

@@

%%% REVIEWER 2 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#2}

{In this work Loos and Romaniello discuss a new implementation of the Bethe-Salpeter equation (BSE) based on the T-matrix approximation. 
The formalism is thoroughly discussed and both a static and dynamical BSE kernel are developed. 
Numerical applications include the calculation of excitation energies of the H2, BeH, and H2O molecules; for the dimers different interatomic distances are considered. 
To establish the accuracy of the new BSE approach based on the T-matrix (BSE@G0T0), full CI reference results are provided. 
Looking at the results it seems that the T-matrix based approaches do not provide a significant improvement with respect to the traditional GW-BSE approach. 
This confirms that in general it is challenging to develop approximations that improve the traditional GW-BSE approach based on the ph-RPA screening. 
Nevertheless, this paper represents a significant effort in going beyond the traditional BSE and could be of interest for the ab initio simulation community and at the base of important future developments. 
By also considering the quality of the presentation and the soundness of the results, I recommend its publication in the Journal of Chemical Physics.}
\\
\alert{
We thank the reviewer for his/her kind comments.
}

\begin{enumerate}

\item 
{The abstract discusses the research that has been carried out, but one or two sentences should be added to describe the outcome/conclusions of this work.}
\\
\alert{
We have modified the abstract to include an additional sentence on the main outcome of the paper: ``Our results suggest that the $T$-matrix-based formalism performs best in few-electron systems where the electron density remains low.''
}

\item 
{In figures 2 and 5 the LUMO@GT curve (diamonds) is completely covered by the LUMO@GW curve (squares). 
This could confuse the reader. 
It would be better, for example, to increase the size of the diamonds to make this curve visible.}
\\
\alert{
We have improved Figs.~2 and 5 and we plot separately the three main quantities (HOMO, LUMO and gap).
}

\item 
{The Green's functions are built from a HF starting point but the authors mention that a self-consistent implementation is available (but not considered in the paper). 
I understand that this implementation might not be fully ready for production, but the authors could at least discuss if they expect improvements from self-consistency or by changing the starting point.}
\\
\alert{
For these small molecular systems, the differences in quasiparticle energies between one-shoot and self-consistent schemes are small.
We have mentioned this observation in the revised version of the manuscript: ``The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here, mainly because, for the small molecular systems studied here (see below), we have observed very small differences between one-shot and self-consistent quasiparticle energies.''
}

### 2022-03-28 Manuscript/Tmatrix.tex
almost done

\begin{abstract}
While the well-established $GW$ approximation corresponds to a resummation of the direct ring diagrams and is particularly well suited for weakly-correlated systems, the $T$-matrix approximation does sum ladder diagrams up to infinity and is supposedly more appropriate in the presence of strong correlation.
Here, we derive and implement, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction are assessed by computing the neutral excited states of molecular systems.
Comparison with more conventional schemes as well as other wave function methods are also reported.
%\bigskip
%\begin{center}
%	\boxed{\includegraphics[width=0.5\linewidth]{TOC}}
%\end{center}
%\bigskip
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} which is freely available at \url{https://github.com/pfloos/QuAcK}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here.
Although the dynamical correction is computed in the dTDA throughout, the zeroth-order excitonic Hamiltonian [see Eq.~\eqref{eq:BSE}] is always the ``full'' BSE static Hamiltonian, i.e., without TDA. 
Reference full configuration interaction (FCI) calculations have been performed with {\QP}. \cite{Garniron_2019}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:ccl}
%%%%%%%%%%%%%%%%%%%%%%
We have derived and implemented, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction have been assessed by computing the neutral excited states of several molecular systems.
Our results suggest that, in the context of the computation of molecular excitation energies, the BSE@$GT$ formalism performs best in few-electron systems where the electron density remains low.
For larger systems where the screening is more important, it seems to be, overall, less accurate than BSE@$GW$.
However, it still outperforms conventional methods such CIS and TDHF for singlet states.

>>>>

\begin{abstract}
While the well-established $GW$ approximation corresponds to a resummation of the direct ring diagrams and is particularly well suited for weakly-correlated systems, the $T$-matrix approximation does sum ladder diagrams up to infinity and is supposedly more appropriate in the presence of strong correlation.
Here, we derive and implement, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction are assessed by computing the neutral excited states of molecular systems.
Comparison with more conventional schemes as well as other wave function methods are also reported.
\alert{Our results suggest that the $T$-matrix-based formalism performs best in few-electron systems where the electron density remains low.}
%\bigskip
%\begin{center}
%	\boxed{\includegraphics[width=0.5\linewidth]{TOC}}
%\end{center}
%\bigskip
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} which is freely available at \url{https://github.com/pfloos/QuAcK}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
\alert{The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here, mainly because, for the small molecular systems studied here (see below), we have observed very small differences between one-shot and self-consistent quasiparticle energies.}
Although the dynamical correction is computed in the dTDA throughout, the zeroth-order excitonic Hamiltonian [see Eq.~\eqref{eq:BSE}] is always the ``full'' BSE static Hamiltonian, i.e., without TDA. 
Reference full configuration interaction (FCI) calculations have been performed with {\QP}. \cite{Garniron_2019}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:ccl}
%%%%%%%%%%%%%%%%%%%%%%
We have derived and implemented, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction have been assessed by computing the neutral excited states of several molecular systems.
Our results suggest that, in the context of the computation of molecular excitation energies, the BSE@$GT$ formalism performs best in few-electron systems where the electron density remains low.
\alert{The overall accuracy of the present scheme still needs to be assessed for larger systems (where the screening is known to be more important). 
For such purposes, a comprehensive benchmark study would be required and we are planning to do so in the future}
%However, it still outperforms conventional methods such CIS and TDHF for singlet states.
